---
title: "NHL Analysis"
author: "Todd Idol"
date: "9/18/2020"
always_allow_html: true
output: 
   github_document:
    toc: true
    toc_depth: 2
---

```{r, setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, warning = FALSE, message = FALSE)
rmarkdown::find_pandoc(version = '2.10.1')
options(knitr.duplicate.label = "allow")
library(tidyverse)
library(jsonlite)
library(httr)
library(rmarkdown)
library(knitr)
```

# Link to Project Repo

Link to repo[https://github.com/tkidol/ST558-Project-1]

# Packages

## Tidyverse

The workhorse of this project for 3 main core packages: DPLYR - manipulated all of my data (selects, joins, filter, new variables... ), Tibble for rendering Data Frames more effectively, & () ggplot2 for discrete & continuous data plots

## Rmarkdown 

Key to the project for the Rmd file itself including the knitr:: functions for consumable object output (kable) & README.md github doc output through render()

## JSONLite

PUsed to converting API pulled JSON data into R objects

## HTTR


Per R documentation site: Functions for the most important http verbs: GET(), HEAD(), PATCH(), PUT(), DELETE() and POST().


# NHL API Wrapper

Created 2 wrapper functions to call APIs.  One for the Franchise endpoints taking in endpoint names, team names & franchise ID's & 1 for the Team stats taking in modifier names, team names & team ID's.

## Franchise Wrapper

```{r, Franchise API Wrapper}
# Franchise Wrapper takes in an API Endpoint (from a list) & Team (either Franchise ID or Common Team name (e.g. "Hurricanes)) & returns distinct API data for analysis.

# Lists to validate incoming arguments
FR_Wrapper <- function(endpoint, team = NULL, ...) {
  menu <- FranchiseAPI()
  records <- c("franchise", "totals", "drilldown", "goalie", "skater")
  teamlist <- list(menu[5])
  frIDs <- list(menu[1])
  
# Validation tests
  if(!(is.character(endpoint))) stop("Invalid Input")
  if(is.character(team)) {
    if ((team %in% teamlist)) stop ("Invalid Input")
  }
   if(is.numeric(team)) {
    if((team %in% frIDs)) stop ("Invalid Input")
  }

  # Conditional logic to pull correct API data
    if(endpoint == "franchise") {
    return(FranchiseAPI())
    
  } else if(endpoint == "totals") { 
     return(TeamTotalsAPI()) 
    
  } else if(endpoint == "drilldown") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Franchise ID`)
      return(TeamDDAPI(team1))
    } else if(is.numeric(team)) {
      return(TeamDDAPI(team))
    }
      
  } else if(endpoint == "goalie") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Franchise ID`)
      return(GoalieAPI(team1))
    } else if(is.numeric(team)) {
      return(GoalieAPI(team))
    }
        
  } else if(endpoint == "skater") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Franchise ID`)
      return(SkaterAPI(team1))
    } else if(is.numeric(team)) {
      return(SkaterAPI(team))
    }
  
  } else {
    return(NULL)
  }
}
```

## Stats Wrapper

```{r, Stat Wrapper}
# Stat API Wrapper takes API Modifier (from a list) & team (Team ID or Common Team name (e.g. "Hurricanes)), & Season (for the 'roster & season' endpoint) & returns distinct API data. Note: using 'people' API for analysis but not directly accessible through Wrapper. MultiAPI works directly with multiple 'id' input but not through the Wrapper

# Lists for comparison to validate incoming arguments
ST_Wrapper <- function(mod, team = NULL, season = 00000000, ...) {
  menu <- FranchiseAPI()
  modifyer <- c("stats", "roster", "person", "nextgame", "priorgame", "expanded", "season")
  teamlist <- list(menu[5])
  teamIDs <- list(menu[4])
 
# Validation tests 
  if((!is.character(mod) & (mod %in% modifyer))) stop("Invalid Input")
  if(is.character(team)) {
      if ((team %in% teamlist)) stop("Invalid Input")
  }
  if(is.numeric(team)) {
    if((team %in% teamIDs)) stop("Invalid Input")
  }
  if(!length(season) != 8) stop("Invalid Input")
  
# Conitional logic to pull correct API data
  if(mod == "stats") {
   return(TeamStatsAPI())
    
  } else if(mod == "roster") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(RosterAPI(team1))
    } else if(is.numeric(team)) {
      return(RosterAPI(team))
    }
      
  } else if(mod == "person") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(PersonAPI(team1))
    } else if(is.numeric(team)) {
      return(PersonAPI(team))
    }
        
  } else if(mod == "nextgame") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(NextGAPI(team1))
    } else if(is.numeric(team)) {
      return(NextGAPI(team))
    }
 
  } else if(mod == "priorgame") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(PriorGAPI(team1))
    } else if(is.numeric(team)) {
      return(PriorGAPI(team))
    }  
    
  } else if(mod == "expanded") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(ExpStatsAPI(team1))
    } else if(is.numeric(team)) {
      return(ExpStatsAPI(team))
    }
  
  } else if(mod == "season") {
    if(is.character(team)) { 
      team1 <- filter(menu,`Name` == team) %>% select(`Team ID`)
      return(SeasonAPI(team1))
    } else if(is.numeric(team, season)) {
      return(SeasonAPI(team, season))
    }  

  } else {
    return(NULL)
  }
}
```


# NHL Franchise APIs

Here are the GET chunks to return APIs associated to the Franchise end points.

### Main API

```{r, Franchise API}
# GET pull for NHL franchise
FranchiseAPI <- function() {
franchise <- GET("https://records.nhl.com/site/api/franchise")

## As JSON text
franchise <- content(franchise, "text")

## As JSON List
franchise <- fromJSON(franchise, flatten = TRUE)

franchise <- franchise[[1]] %>% rename("Franchise ID" = id, "First Season" = firstSeasonId, "Last Season" = lastSeasonId, "Team ID" = mostRecentTeamId, "Name" = teamCommonName, "Location" = teamPlaceName)
return(franchise)
}
```

### Wrapper Test
```{r, Franchise API Head, echo = FALSE}
# API head
frhead <- FR_Wrapper("franchise")
kable(head(frhead))
```

### Team Totals API

```{r TeamTotal API}
# GET pull for NHL Team Totals
TeamTotalsAPI<- function() {
totals <- GET("https://records.nhl.com/site/api/franchise-team-totals")

## As JSON text
totals <- content(totals, "text")

## As JSON List
totals <- fromJSON(totals, flatten = TRUE)
return(totals[[1]])
}
```

```{r, Totals API Head, echo = FALSE}
# API head
tothead <- FR_Wrapper("totals")
kable(head(tothead))
```

### Team Drill Down API

```{r, Team DrillDown API}
TeamDDAPI <- function(id, ...) {
# GET pull for NHL Team Totals
drill_down <- GET(paste0("https://records.nhl.com/site/api/franchise-season-records?cayenneExp=franchiseId=", id))

## As JSON text
drill_down <- content(drill_down, "text")

## As JSON List
drill_down <- fromJSON(drill_down, flatten = TRUE)
return(drill_down[[1]])
}
```
### Wrapper test

```{r, Drilldown API Head, echo = FALSE}
# API head
ddhead <- FR_Wrapper("drilldown", 26)
kable(head(ddhead))
```

### Goalie API

```{r Goalie API}
GoalieAPI <- function(id, ...) {
# GET pull for NHL Team Totals
goalie <- GET(paste0("https://records.nhl.com/site/api/franchise-goalie-records?cayenneExp=franchiseId=", id))

## As JSON text
goalie <- content(goalie, "text")

## As JSON List
goalie <- fromJSON(goalie, flatten = TRUE)
return(goalie[[1]])
}
```
### Wrapper Test

```{r, Goalie API Head, echo = FALSE}
# API head
ghead <- FR_Wrapper("goalie", "Avalanche")
kable(head(ghead))
```

### Skater API

```{r, Skater API}
SkaterAPI <- function(id, ...) {
# GET pull for NHL Team Totals
skater <- GET(paste0("https://records.nhl.com/site/api/franchise-skater-records?cayenneExp=franchiseId=", id))

## As JSON text
skater <- content(skater, "text")

## As JSON List
skater <- fromJSON(skater, flatten = TRUE)
return(skater[[1]])
}
```
### Wrapper Test

```{r, Skater API Head, echo = FALSE}
# API head
skhead <- FR_Wrapper("skater", 2)
kable(head(skhead))
```

### Complete Team Stats

```{r, Team Stats API}
TeamStatsAPI <- function() {
stats <- GET("https://statsapi.web.nhl.com/api/v1/teams")

## As JSON text
stats <- content(stats, "text")

## As JSON List
stats <- fromJSON(stats, flatten = TRUE)
return(stats[[2]])
}
```

### Wrapper Test

```{r, Team ST API Head, echo = FALSE}
# API head
tshead <- ST_Wrapper("stats")
kable(head(tshead))
```

### Roster API

```{r, Roster API}
RosterAPI <- function(id, ...) {
roster <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "/?expand=team.roster"))

## As JSON text
roster <- content(roster, "text")

## As JSON List
roster <- fromJSON(roster, flatten = TRUE)
franchiseId <- roster$teams$franchiseId
roster <- as_tibble(roster[[2]]$roster.roster[[1]]) %>% mutate("franchiseId" = franchiseId) %>% 
select(franchiseId, everything())
return(roster)
}
```

### Wrapper Test

```{r, Roster Test, echo = FALSE}
# API head
roshead <- ST_Wrapper("roster", 15)
kable(head(roshead))
```

### Person API

```{r, Person API}
PersonAPI <- function(id, ...) {
person <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "/?expand=person.names"))

## As JSON text
person <- content(person, "text")

## As JSON List
person <- fromJSON(person, flatten = TRUE)
return(person[[2]])
}
```

### Wrapper Test
```{r, Person API Head, echo = FALSE}
# API head
pershead <- ST_Wrapper("person", "Sharks")
kable(head(pershead))
```
### People API (not in wrapper - internal use)

```{r, People API}
# Detailed substitue for PersonAPI
PeopleAPI <- function(id, ...) {
people <- GET(paste0("https://statsapi.web.nhl.com/api/v1/people/", id))

## As JSON text
people <- content(people, "text")

## As JSON List
people <- fromJSON(people, flatten = TRUE)
return(people[[2]])
}
```

### Team Player Function (for internal analysis use)

```{r, team player}
## New DF combining People with Roster for additional player stats
TeamPlayer <- function (id, ...) {
  
# Get player ID from Roster
playerId <- pull(RosterAPI(id), person.id)

# Seed data frame
players <- as.data.frame(PeopleAPI(playerId[1]))

# Iterate player IDs over People API to build data frame 
  for (i in 2:length(playerId)) {
    players[i, ] <- (PeopleAPI(playerId[i]))
  }

# Trim for relevant data for analysis
players <- players %>% select(currentTeam.id, id, fullName, birthCountry, primaryPosition.name)
if(length(players$currentTeam.id) > 2) {
  players$currentTeam.id = players$currentTeam.id[[1]]
}
return(players) 
}
```
### Next Game API

```{r, Next Game API}
NextGAPI <- function(id, ...) {
nextG <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "?expand=team.schedule.next"))

## As JSON text
nextG <- content(nextG, "text")

## As JSON List
nextG <- fromJSON(nextG, flatten = TRUE)
nextG <- nextG[[2]]$nextGameSchedule.dates[[1]]$games[[1]]
return(nextG)
}
```

### Wrapper Test

```{r, NG API Head, echo = FALSE}
# API head
nghead <- ST_Wrapper("nextgame", "Lightning")
kable(head(nghead))
```

### Prior Game API

```{r, Prior Game API}
PriorGAPI <- function(id, ...) {
priorG <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "?expand=team.schedule.previous"))

## As JSON text
priorG <- content(priorG, "text")

## As JSON List
priorG <- fromJSON(priorG, flatten = TRUE)
priorG <- priorG[[2]]$previousGameSchedule.dates
return(priorG)
}
```

### Wrapper Test

```{r, Prior API Head, echo = FALSE}
# API head
pghead <- ST_Wrapper("priorgame", 28)
kable(head(pghead))
```

### Expanded Stats API

```{r, Expanded Stats API}
ExpStatsAPI <- function(id, ...) {
expStats <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "?expand=team.stats"))

## As JSON text
expStats <- content(expStats, "text")

## As JSON List
expStats <- fromJSON(expStats, flatten = TRUE)
expStats <- as.data.frame(expStats[[2]]$teamStats[[1]]$splits)
return(expStats)
}
```
### Wrapper Test

```{r, Exp API Head, echo = FALSE}
# API head
exhead <- ST_Wrapper("expanded", 28)
kable(head(exhead))
```
### Season API (tested in Data Collection)

```{r, Season Stats API}
SeasonAPI <- function(id, season, ...) {
season <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "?expand=team.roster&season=", season))

## As JSON text
season <- content(season, "text")

## As JSON List
season <- fromJSON(season, flatten = TRUE)
franchiseId <- season$teams$franchiseId
teamName <- season$teams$teamName
season <- as_tibble((season[[2]]$roster.roster[[1]])) %>% mutate("franchiseId" = franchiseId, "teamName" = teamName) %>% select(franchiseId, everything())
return(season)
}
```
### Multi Team API (no wrapper but functions directly)

```{r, Multi Teams API}
MultiAPI <- function(id, ...) {
multi <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", "?teamId=", id))

## As JSON text
multi <- content(multi, "text")

## As JSON List
multi <- fromJSON(multi, flatten = TRUE)
return(multi[[2]])
}
```
### Playoff API (void of valuable info)

```{r, Playoffs API}
PlayoffAPI <- function(id, ...) {
playoff <- GET(paste0("https://statsapi.web.nhl.com/api/v1/teams/", id, "?stats=statsSingleSeasonPlayoffs"))

## As JSON text
playoff <- content(playoff, "text")

## As JSON List
playoff <- fromJSON(playoff, flatten = TRUE)
return(playoff[[2]])
}
```

# 2017 - 2019 Presidents Cup Teams vs. Low Points Teams Analysis



## Data

The NHL team with the highest regular season points (2 for ea win / 1 for ea overtime loss) is awarded the league's Presidents Cup.  The data used for the analysis is taken for a comparison of the 2017 - 2019 Presidents Cup Winners (Capitals, Predators, Lightning) vs the corresponding "Low Points Teams" (Avalanche, Sabres, Senators).  SeasonAPI was used to pull Player info for each of the six teams for corresponding years.  This was mutated & joined to the PeopleAPI data to get nationalities for each Player (factored into USA, Canada, European groups).  That data was then joined to the SkaterAPI to get career stats on which the Rosters were analyzed. 

```{r, presidents, cache = TRUE}
# Roster for last 2 seasons, high points team (presidents cup) low points team
pres17 <- SeasonAPI(15, 20162017)
last17 <- SeasonAPI(21, 20162017)
pres18 <- SeasonAPI(18, 20172018)
last18 <- SeasonAPI(7, 20172018)
pres19 <- SeasonAPI(14, 20182019)
last19 <- SeasonAPI(9, 20182019)

# Combine results into single DS & add year & finish
PresLast <- bind_rows(pres17, last17, pres18, last18, pres19, last19)
PresLast <- mutate(PresLast, year = 
                  ifelse((PresLast$teamName == "Capitals") | (PresLast$teamName == "Avalanche"), 17,
                  ifelse((PresLast$teamName == "Predators") | (PresLast$teamName == "Sabres"), 18, 19)))
PresLast <- mutate(PresLast, rank = ifelse((PresLast$teamName == "Capitals") | (PresLast$teamName == "Predators") | (PresLast$teamName == "Lightning"), "First", "Last"))
PresLast <- PresLast %>% select(franchiseId, teamName, year, rank, person.id, person.fullName, jerseyNumber, position.code)


# Franchise ID's for SkaterAPI argument
frIdList <- c(pres17$franchiseId[1], last17$franchiseId[1], pres18$franchiseId[1],last18$franchiseId[1], pres19$franchiseId[1], last19$franchiseId[1])

# Skater for detailed stats
PresLast_Skater <- lapply(frIdList, SkaterAPI)
PresLast_Skater <- bind_rows(PresLast_Skater)

# Combine results into single DS
PresLast_Skater <- PresLast_Skater %>% rename("person.id" = playerId) 

# Join Preslast & Preslast_Skater fo Skaters on Pres-last Teams / Seasons for analysis
PL_SK_Season <- inner_join(PresLast, PresLast_Skater) 
PL_SK_Season <-as_tibble(PL_SK_Season) %>% select(franchiseId, teamName, year, rank, person.id, person.fullName, jerseyNumber, position.code, seasons, penaltyMinutes, goals, assists, mostPointsOneSeason) %>% distinct()

# Get additional categorical stats for Players
Season_People <- lapply(PL_SK_Season$person.id, PeopleAPI)
Season_People <- bind_rows(Season_People) %>% select(id, nationality) %>% mutate("person.id" = id)

# Create regions for categorical analysis
PL_SK_Season <- inner_join(PL_SK_Season, Season_People)
PL_SK_Season <- mutate(PL_SK_Season, country = ifelse((nationality == "USA"), "USA",
                                               ifelse((nationality == "CAN"), "CAN", "EUR")))

# Create Average PM & Points stats for Players
PL_SK_Season <- PL_SK_Season %>% mutate(AveragePM = round((penaltyMinutes/seasons), 2), TotalPoints = (goals+assists), AverageTP = round((goals+assists)/seasons), 2)

# Select relevant variables for analysis
PL_SK_Season <- PL_SK_Season %>% select(franchiseId, teamName, year, rank, person.fullName, country, position.code, seasons, AveragePM, AverageTP) %>% distinct()

# Rename for Graphical & Table prentation
PL_SK_Season <- PL_SK_Season %>% rename("Franchise ID" = franchiseId, "Team" = teamName, "Year" = year, "Finish" = rank, "Name" = person.fullName, "Country" = country, "Position" = position.code, "Player Seasons" = seasons, "Avg Penalty Min" = AveragePM, "Avg Total Pts" = AverageTP)

# Create factor for Team Finish
PL_SK_Season$Finish <- as.factor(PL_SK_Season$Finish)

# Print sample table for report
kable(head(PL_SK_Season), caption = "Preview of 2017 - 2019 Team Rosters: Presidents Cup vs Last Place")
```

# Categorical Analysis

Categories for Player Countries (regionalized), Positions (by code), & Team Finish.  

## Total Summaries



```{r, Team Finish }
Pos_Ctry_table <- table(PL_SK_Season$Position, PL_SK_Season$Country)
kable(Pos_Ctry_table, caption = "Player Position by Country")
```

### Summaries by First Place Teams

```{r, Pres PM_TP Rank}
First <- filter(PL_SK_Season, PL_SK_Season$Finish == "First")
Pos_Ctry1_table <- table(First$Position, First$Country)
kable(Pos_Ctry1_table, caption = "Presidents Cup Winners")
```

### Summaries by Last Place Teams

```{r, Last PM_TP Rank}
Last <- filter(PL_SK_Season, PL_SK_Season$Finish == "Last")
Pos_Ctry2_table <- table(Last$Position, Last$Country)
kable(Pos_Ctry2_table, caption = "Last Place Players")
```
## Categorical Visuals

### 1 Factor

The most interesting finding from the single category plot is the smaller roster size of the 3 First Place teams.  This could indicate fewer injuries or the fact that the Last Place teams were juggling lineups and bringing in players from their minor league ranks or through trades.

```{r, univariate bar graphs}
# bar plot of Team Finish
FinPlot <- ggplot(PL_SK_Season, aes(x = Finish), color = "blue") 
FinPlot + geom_bar() + labs(x = "Player Finish")

# bar plot of Team Finish
COplot <- ggplot(PL_SK_Season, aes(x = Country), color = "blue") 
COplot + geom_bar() + labs(x = "Player Country")

# bar plot of Team Finish
COPlot <- ggplot(PL_SK_Season, aes(x = Position), color = "blue") 
COPlot + geom_bar() + labs(x = "Player Position")
```

### 2 Factors

Consistent with univariate analysis we see fewer players on the winning teams.  It's notable that Canadians dominate both rosters (they invented the sport).  Europeans and Americans are approximately the same.  Centers and Defensemen are clearly the most valuable positions to fill - strong stats in these groups could show significant differences between First / Last groups in the Qantitative Analysis. 

```{r, 2 factor bar graphs}
# Finish by Nationality
FxCOplot <- ggplot(PL_SK_Season, aes(Finish, fill = as.factor(Country)))
FxCOplot+ geom_bar(position = "dodge") +  scale_fill_discrete(name = "Country") + labs(x = "Team Finish")

# Country by Position
FxPosplot <- ggplot(PL_SK_Season, aes(Country, fill = as.factor(Position)))
FxPosplot + geom_bar(position = "dodge") +  scale_fill_discrete(name = "Position") + labs(x = "Country Region")
```


### 3 Factors

Basically a breakdown of the previous 2 bivariate plots demonstrating the same conclusions.  The number of Canadian defensemen on Last Place rosters stands out - perhaps a crop of younger players.

```{r, 3 factor bar graphs}
# Color by satellite indicator by spine condition
FxPMxPtplot <- ggplot(PL_SK_Season, aes(Country, fill = Finish))
FxPMxPtplot + geom_bar(position = "dodge") + scale_fill_discrete(name = "Team Finish") + labs(x = "Player Country", title = "Player Finish by Country and Position") + facet_wrap(. ~ Position, labeller = (label_both))
```

# Quantitative Analysis

## Summaries

```{r, player summaries}
# Function to knit 5 number sum + Mean given Species input
TeamSumm <- function(team, ...) {
summ <- PL_SK_Season %>% filter(Team == team) %>% select(c(8, 9, 10)) %>% apply(2, summary)
kable(round(summ, 2), caption = paste0("Team: ", team))
}
```

```{r, TeamStats call }
teams <- (c("Capitals", "Avalanche", "Predators", "Sabres", "Lightning", "Senators"))

# Capitals summary
TeamSumm(teams[[1]])

# Avalanche summary
TeamSumm(teams[[2]])

# Predators summary
TeamSumm(teams[[3]])

# Sabres summary
TeamSumm(teams[[4]])

# lightning summary
TeamSumm(teams[[5]])

# Senators summary
TeamSumm(teams[[6]])

```

## Visuals

### Average Total Points vs Player Tenure

An apparent strong correlation between these variables especially towards the upper right indicating veteran players produce more points (goals + assists) on average over their careers.  This could be the result of their skill leading to longevity vs experience producing more points (interesting topic for deeper analysis).  Wide variance given by the large CI bands.  

```{r, dotplot by TotalPoints}
# Base plot aesthetic with Total Points on x axis
Finish <- PL_SK_Season$Finish
Avg_Total_Points <- PL_SK_Season$`Avg Total Pts`
Avg_Penalty_Min <- PL_SK_Season$`Avg Penalty Min`
PLayer_Seasons <- PL_SK_Season$`Player Seasons`

g <- ggplot(PL_SK_Season, aes(x = Avg_Total_Points, y = PLayer_Seasons, color = Finish))

# Avg total points histogram
g + geom_point() + geom_smooth(aes(group = Team), method = lm) + scale_fill_continuous() + labs(title =  "Players Average Total Points vs Seasons")
```

### Penalty Mins by Player Tenure

Mixed results among First & Last Placed teams but generally First Place teams show fewer PMs by season.  Could indicate "maturity" and recognition of importance of "staying out of the box" to winning over a long season.

```{r, dotplot by Penalty Minutes}
# Base plot aesthetic with Total Points on x axis
g <- ggplot(PL_SK_Season, aes(x = Avg_Penalty_Min, y = PLayer_Seasons, color = Finish))

# Avg total points histogram
g + geom_point() + geom_smooth(aes(group = Team), method = lm) + scale_fill_continuous() + labs(title =  "Players Average Penalty Minutes vs Average Total Points")
```

### Average Total Points vs Average Penalty Minutes

Breaking out colors by team to show another degree of team detail.  Tamapa Bay jumps out here for the steep Beta 1 - players getting the most points out of their penalty minutes & shortness of line indicate fewer penalty minute outliers (consistent blend of aggression & discipline across roster).  Capitals are next highest in Career points per PM but w less discipline (aggressive play leading to positive results). The 3 Last Place teams stand out for the shallowness (low points per penalty mins) of their lines & clustering in the lower left (fewer points/fewer penalty minutes) - teams appear to be less aggressive/offensively minded - could be Experience related.  The 3 Last Place teams have man tenures under 4 years while the First Place teams are over 5 as a group.  Capitals have the max tenured player at 15 years - The Great 8, Alex Ovechkin - you'll see him at the top of all charts. He's pretty good.

```{r, dotplot by }
# Base plot aesthetic with Total Points on x axis
g <- ggplot(PL_SK_Season, aes(x = Avg_Penalty_Min, y = Avg_Total_Points, color = Team))

# Avg total points histogram
g + geom_point() + geom_smooth(aes(group = Team), method = lm) + scale_fill_continuous() + labs(title =  "Players Average Total Points vs Average Penalty Minutes")
```


